### 2024-05-28 fantasy-football.qmd
20240527 - glossary

- Quarterback (QB)
- Running Back (RB)
    - Halfback or Tailback
    - Fullback (FB)
- Wide Receiver (WR)
- Tight End (TE)
- Offensive Linemen (OL)
    - Center (C)
    - Offensive Guard (OG)
    - Offensive Tackle (OT)

Quarterbacks are the most important player on the offense.
They help lead the team down the field.
Quarterbacks receive the ball from the Center at the beginning of the play, and they can either hand the ball off (typically to a Running Back or Fullback), pass the ball (typically to a Wide Receiver or Tight End), or run the ball.
Quarterbacks tend to have a strong arm for throwing the ball far and accurately.
Some quarterbacks are fast and are considered "dual threats" to pass or run.

Running Backs take a hand-off from the Quarterback to execute a running play.
They may also catch short passes from the Quarterback or help protect (i.e., block for) the Quarterback from the defensive players who are trying to tackle the Quarterback.
Halfbacks or Tailbacks tend to be quick and agile.
Fullbacks tend to be strong and powerful.

@@

- Defensive Linemen (DL)
    - Defensive End (DE)
    - Defensive Tackle (DT)
- Linebacker (LB)
    - Middle (or Inside) Linebacker (MLB)
    - Outside Linebacker (OLB)
- "The Secondary" / Defensive Back (DB)
    - Cornerback (CB)
    - Safety (S)
        - Free Safety (FS)
        - Strong Safety (SS)

@@

On passing plays, Defensive Linemen try to apply pressure to the Quarterback and try to sack them.
On rushing plays, Defensive Linemen try to tackle the ball carrier.
Defensive Linemen tend to be large yet quick so they can apply pressure to the Quarterback.

@@

The football field is rectangular and is 120 yards long and 53 1/3 yards wide.
At each end of the 120-yard field is a team's end zone.
Each end zone is 10 yards long.
Thus, the distance from one end zone to the other end zone is 100 yards.
Behind each end zone is a field goal post.
A diagram of a football field is depicted in Figure @fig-footballField.

@@

The team with the ball (i.e., the offense) has four opportunities ("downs") to advance the ball 10 yards.
A team can advance the ball either by running it or passing and catching it.
At the end of a rushing play, the ball advances to wherever the ball carrier is tackled or goes out of bounds (i.e., wherever the player is "down").
At the end of a passing play, if the thrown ball is caught (i.e., a completed pass), the ball advances to wherever the ball carrier is tackled or goes out of bounds.
If the thrown ball is not caught in bounds before the ball hits the ground (i.e., an incomplete pass), the ball does not advance.
Wherever the ball is advanced to dictates where the next play begins.
The yard position on the field where the next play takes place from is known as the "line of scrimmage".
Neither team can cross the line the line of scrimmage until the next play begins.
To begin the play, the ball is placed on the line of scrimmage and the Center gives (or "snaps") the ball to the Quarterback.

@@

There are multiple ways that ball possession can switch from the offense to the other team.
After scoring a touchdown, field goal, or safety, there is a kickoff, in which the scoring team kicks the ball to the opponent.
Another way that the ball swithces possession to the other team is if the team commits a turnover.
The defense can force a turnover by an interception, fumble recovery, or turnover on downs.
A turnover due to an interception occurs when the defense catches the the quarterback's pass.
A turnover due to a fumble recovery occurs when an offensive player, who had possession of the ball, loses the ball before being down or scoring a touchdown and the ball is recovered by the opponent.
A turnover on downs occurs when the team attempts to achieve the remainder of the needed 10 yards to go on fourth down but fails.

@@

The goal of the team with the ball (i.e., the offense) is to score points.
It can do this by either advancing the ball into the other team's end zone (6 points) or by kicking a field goal (3 points).
Advancing the ball in the other team's end zone is called a touchdown.
After a touchdown, the offense chooses to attempt either a point-after-touchdown (PAT) or a two-point conversion.
A PAT is a short kick attempt from the 15-yard line that, if it goes through the goal posts ("uprights"), is worth 1 point.
A two-point conversion is a single-scoring opportunity from the 3-yard line (i.e., 3 yards away from the end zone).
If the offense scores (i.e., advances the ball into the end zone) from the 3-yard line, the team is awarded 2 points.

@@

A safety occurs when the offense is tackled with the ball in their own end zone.
When a safety occurs, the opposing team (i.e., defense) is awarded two points.

@@

- running play ("run") or rushing play (or "rush")
- passing play (or "pass")
- passing attempt
- rushing attempt
- passing completion
- passing incompletion
- passing yards
- rushing yards
- receving yards
- reception
- touchdown
- passing touchdown
- rushing touchdown
- receiving touchdown
- two-point conversion
- block
- kickoff
- field goal
- point after touchdown (PAT)
- extra point returned
- punt
- fumble lost
- fumble forced
- fumble recovery
- interception
- tackle
- tackle solo
- tackle assist
- tackle for loss
- sack
- pass defended
- safety

>>>>

- Quarterback (QB)
- Running Back (RB)
    - Halfback (HB) or Tailback (TB)
    - Fullback (FB)
- Wide Receiver (WR)
- Tight End (TE)
- Offensive Linemen (OL), part of the "Offensive Line"
    - Center (C)
    - Offensive Guard (OG)
    - Offensive Tackle (OT)

The quarterback is the most important player on the offense.
They help lead the team down the field.
The quarterback receives the ball from the Center at the beginning of the play, and they can either hand the ball off (typically to a Running Back or Fullback), pass the ball (typically to a Wide Receiver or Tight End), or run the ball.
Quarterbacks tend to have a strong arm for throwing the ball far and accurately.
Some quarterbacks are fast and are considered "dual threats" to pass or run.

Running Backs take a hand-off from the Quarterback to execute a running play (i.e., a rush).
They may also catch short passes from the Quarterback or help protect (i.e., block for) the Quarterback from the defensive players who are trying to tackle the Quarterback.
Halfbacks and Tailbacks tend to be quick and agile.
Fullbacks tend to be strong and powerful.

@@

- Defensive Linemen (DL), part of the "Defensive Line"
    - Defensive End (DE)
    - Defensive Tackle (DT)
- Linebacker (LB)
    - Middle (or Inside) Linebacker (MLB)
    - Outside Linebacker (OLB)
- Defensive Back (DB), part of the "Secondary"
    - Cornerback (CB)
    - Safety (S)
        - Free Safety (FS)
        - Strong Safety (SS)

@@

On passing plays, Defensive Linemen try to apply pressure to the Quarterback and try to tackle the Quarterback behind the line of scrimmage before the Quarterback can throw the ball (i.e., a sack).
On rushing plays, Defensive Linemen try to tackle the ball carrier to prevent the ball carrier from advancing the ball (i.e., gaining yards).
Defensive Linemen tend to be large yet quick so they can apply pressure to the Quarterback.

@@

The football field is rectangular and is 120 yards long and 53 1/3 yards wide (109.73 m x 48.77 m).^[One yard is equal to three feet. A yard is just smaller than a meter (0.9144 meters).]
At each end of the 120-yard field is a team's end zone.
Each end zone is 10 yards long (9.14 m).
Thus, the distance from one end zone to the other end zone is 100 yards (91.44 m).
Behind each end zone is a field goal post.
A diagram of a football field is depicted in Figure @fig-footballField.

@@

The team with the ball (i.e., the offense) has four opportunities ("downs") to advance the ball (i.e., gain) 10 yards.
A team can advance the ball either by running it or by throwing (i.e., passing) and catching it.
At the end of a rushing play, the ball advances to wherever the ball carrier is tackled or goes out of bounds (i.e., wherever the player is "down").
At the end of a passing play, if the thrown ball is caught (i.e., a completed pass), the ball advances to wherever the ball carrier is tackled or goes out of bounds.
If the thrown ball is not caught in bounds before the ball hits the ground (i.e., an incomplete pass), the ball does not advance.
Wherever the ball is advanced to dictates where the next play begins.
The yard position on the field where the next play takes place from is known as the "line of scrimmage".
Neither team can cross the line the line of scrimmage until the next play begins.
To begin the play, the ball is placed on the line of scrimmage and the Center gives (or "snaps") the ball to the Quarterback.

@@

There are multiple ways that ball possession can switch from the offense to the other team.
After scoring a touchdown, field goal, or safety, there is a kickoff, in which the scoring team kicks the ball to the opponent.
Another way that the ball switches possession to the other team is if the team commits a turnover.
The defense can force a turnover by an interception, fumble recovery, or turnover on downs.
A turnover due to an interception occurs when a defensive player catches the quarterback's pass.
A turnover due to a fumble recovery occurs when an offensive player, who had possession of the ball, loses the ball before being down or scoring a touchdown and the ball is recovered by the opponent.
A turnover on downs occurs when the team attempts on fourth down to achieve the remainder of the needed 10 yards to go but fails.

@@

The goal of the team with the ball (i.e., the offense) is to score points.
It can do this by either advancing the ball into the other team's end zone (6 points) or by kicking a field goal (3 points).
Advancing the ball in the other team's end zone is called a touchdown.
After a touchdown, the offense chooses to attempt either a point-after-touchdown (PAT) or a two-point conversion.
A PAT is a short kick attempt from the 15-yard line (i.e., 15 yards away from the end zone) that, if it goes through the goal posts ("uprights") and over the cross bar, is worth 1 point.
A two-point conversion is a single-scoring opportunity from the 3-yard line (i.e., 3 yards away from the end zone).
If the offense scores (i.e., advances the ball into the end zone) from the 3-yard line, the team is awarded 2 points.

@@

A safety occurs when the offense is tackled with the ball in their own end zone.
When a safety occurs, the opposing team (i.e., defense) is awarded two points and the ball.

@@

- running play ("run") or rushing play (or "rush")—the attempt by an offensive player, typically the Running Back or Quarterback, to advance the ball "on the ground" by running it—not by passing it forward
- passing play (or "pass")—the attempt by an offensive player, typically the Quarterback, to advance the ball by throwing it forward to an offensive player
- passing attempt—the attempt to advance the ball by passing it (i.e., a thrown pass)
- rushing attempt—the attempt to advance the ball by running it
- passing completion—a thrown pass that is succesfully caught by an offensive player
- passing incompletion—a thrown pass that is not caught by an offensive player
- passing yards—the distance (in yards) the player advanced the ball by throwing it
- rushing yards—the distance (in yards) the player advanced the ball by running it
- receving yards—the distance (in yards) the player advanced the ball by catching thrown passes and then running with it further upfield
- kick/punt return yards—the distance (in yards) the player advanced the ball by returning kicks or punts
- turnover return yards—the distance (in yards) the player advanced the ball by returning turnovers
- reception—a pass that is caught by the offensive player
- touchdown—advancing the ball into the opponent's end zone either by a) throwing a completed pass that ends up in the end zone, b) running it into the end zone, c) catching it in the end zone, or d) catching it and then running it into the end zone
- passing touchdown—advancing the ball into the opponent's end zone either by throwing a completed pass that ends up in the end zone
- rushing touchdown—advancing the ball into the opponent's end zone either by running it into the end zone
- receiving touchdown—advancing the ball into the opponent's end zone either by catching it in the end zone or by catching it and then running it into the end zone
- kick/punt return touchdown—advancing the ball into the opponent's end zone when returning a kick or punt
- turnover return touchdown—advancing the ball into the opponent's end zone when returning a turnover (i.e., interception or fumble)
- two-point conversion—a single-scoring opportunity from the 3-yard line (i.e., 3 yards away from the end zone) that is an option given to a team that scores a touchdown;
if the offense scores (i.e., advances the ball into the end zone) from the 3-yard line, the team is awarded 2 points
- block—when the defense/special teams blocks a kick or field goal by hitting the ball just after it is kicked to prevent the ball from going far
- kickoff—the kicking team kicks the ball to the receiving team, who has the option to return the kick
- field goal—a kicker kicks the ball with an intent to kick the ball through the field goal posts ("uprights").
To score points by making a field goal, the kicked ball must go between the uprights (extended vertically) and over the cross bar.
If the field goal attempt is successful, the team gains 3 points.
- point after touchdown (PAT)—a short kick attempt from the 15-yard line (i.e., 15 yards away from the end zone) that, if it goes through the goal posts ("uprights") and over the cross bar, is worth 1 point
- extra point returned—if the defense/special teams returns the ball into the opponent's end zone during a point after touchdown (PAT) attempt, it is worth 2 points
- punt—a punter kicks the ball to the other team with an intent to give their opponent worse field position, thus making it harder for the other team to score
- fumble lost—when an offensive player, who had possession of the ball, loses the ball before being down or scoring a touchdown and the ball is recovered by the opponent
- fumble forced—when a defensive player knocks the ball out of the hands of an offensive player, who had possession of the ball
- fumble recovery—when a defensive player recovers a fumble by the opponent
- interception—when a defensive player catches a pass from an offensive player
- tackle—when a player brings down the ball carrier
- tackle solo—when a player is the main tackler (i.e., the primary player to bring down the ball carrier)
- tackle assist—when a player is one of two or more players who, together, bring down the ball carrier
- tackle for loss—when an offensive player is tackled with the ball behind the line of scrimmage
- sack—when a Quarterback is tackled with the ball behind the line of scrimmage
- pass defended—when a defensive player knocks down the ball in the air so that the indended receiver cannot catch the ball
- safety—when the offense is tackled with the ball in their own end zone
